### 2023-02-06 survey_results.Rmd
co-authorship networks

---
title: "Survey results"
author: "Juan Rocha"
date: April 2022
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r}
library(tidyverse)
load("data/survey_data.RData")
```

The survey ran between `r range(dat$StartDate |> lubridate::as_date()) |> lubridate::parse_date_time("y m d")`. `r nrow(dat)` people answered the survey, of which `r dat |> filter(Q4 == "Yes") |> nrow()` gave us consent to use the results internally at SRC. All results below include only the individuals who gave us their permission.