### 2025-07-28 review_paper/outline.Rmd
back up old stuff

---
title: "Extended outline"
date: "2024-08-15"
output:
  pdf_document:
    dev: pdf
    keep_tex: yes
    toc: no
    latex_engine: xelatex
csl: /Users/juanrocha/Documents/styles/annual-reviews.csl
bibliography: refs.bib
citecolor: blue
documentclass: article
---

## ***Title:*** Regime shifts research meets transformations research - advancing critical frontiers for safe and just futures

### ***Key insights** (to be updated):*

-   Clarify the difference and overlaps between transformations and regime shifts theories and the use of concepts such as tipping points
-   Highlighting frontiers of research that need both approaches
-   Navigating the space of positive and negative tipping points

### Literature cited

We will review literature from the social and natural sciences regarding regime shifts and transformations research.
These two fields are typically dealt with in natural resource management, environmental economics, ecology, sustainability science, rural studies, political ecology, among others.
Examples of seminal work on these topics include Holling [-@holling1973resilience], Folke et al [-@folke2004], Olsson et al [-@olsson2006shooting], or books such as Navigating Social-Ecological systems [@berkes2008navigating], Linking social ecological systems [@berkes2000linking], or Critical transitions in Nature and Society [@scheffer2009critical].
Below you find the outline of our review with sample references of the work we are reviewing.

## Outline

-   **Problem setting:** There are two strands of literature within the field of sustainability science and natural resource management that share a common origins [@holling1973resilience; @gunderson2002panarchy; @folke2004; @folke2010resilience; @folke2005adaptive] but have been diverging in the last couple of decades: regime shifts and transformations research.
    In diverging, there are risks for oversimplification and misuse of concepts, jargonification, using multiple words to mean the same thing, but importantly also the lack of cross-fertilization between the two different scholarships.

-   **Definitions:** what are regime shifts, what are transformations ?[
    1000w, 500w each]

    -   Regime shifts key characteristics: nonlinear change characterised by tipping points.
        Important features include hysteresis, feedback mechanisms, with potential impacts on ecosystem services and human wellbeing [@lewontin1969meaning; @holling1973resilience; @scheffer2001; @beisner2003alternative; @scheffer2003].

    -   Transformations key characteristics: nonlinear changes characterised by shifting basins of attractions towards more desirable states.
        Important features include windows of opportunity, agency and power dynamics [@gunderson2002panarchy; @walker2004resilience; @olsson2006shooting; @folke2010resilience; @herrfahrdt-pähle2020; @geels2018socio; @olsson2024transformations].

-   **History of both streams of research** (nice graphics, iconic papers) [1000w]

    -   Timelines

        -   Regime shifts: [@lewontin1969meaning; @holling1973resilience; @scheffer2001; @beisner2003alternative; @scheffer2003; @carpenter2005; @andersen2009; @carpenter2009; @sole2011phase; @lade2015; @rocha2018; @martin2020]

        -   Transformations: [@gunderson2002panarchy; @walker2004resilience; @westley2011tipping; @olsson2006shooting; @folke2010resilience; @moore2012loop; @moore2014studying; @avelino2017power; @geels2018socio; @mcgowan2023recognizing; @herrfahrdt-pähle2020; @moore2020imagination; @olsson2024transformations]

    -   Initial focus of both fields and and how they changed over time, including highlighting phases of theory vs. empirical focus of both fields and applied methods.
        This section will also include work that put forward skeptic points of view on the validity of regime shifts or transformation theories (e.g. [@hillebrand2020; @mcgowan2023recognizing] )

-   **Commonalities and differences** (e.g. power in transformations research vs hysteresis in regime shifts research) [500w]

    -   What questions do they ask? What methods do they use? What disciplinary approaches are involved (targeted)? What scale? What are the main findings?

-   **Current research challenges** [1500w, 500w each]

    -   Multidimensionality

    -   Observability and detection

    -   Management: ethical stakes

-   **Navigating the space of positive and negative tipping points** [1000w, 500w each]:

    -   Criticism of current (mis)uses of positive and negative tipping points for their social desirability (e.g. [@lenton2022operationalising; @lenton2023global])

    -   Distribution of costs, benefits, and power to make a difference

    -   Risk of reinforcing colonial practices [@pereira2024equity]

-   **Conclusions**: frontiers of research

### Example of literature reviewed