### 2025-07-29 review_paper/manuscript.Rmd
intro to RS + fixing template

---
title: "Regime shifts and transformations: "
subtitle: "Advancing critical frontiers for safe and just futures"
author: |
    | Juan C. Rocha et al.$^1$ 
    | \affil{$^1$Stockholm Resilience Centre, Stockholm University, Stockholm, Sweden 10691; email: \texttt{\href{mailto:juan.rocha@su.se}{\nolinkurl{juan.rocha@su.se}}}}

output:
  pdf_document:
    keep_tex: yes
    keep_md: yes
    latex_engine: xelatex
    #template: template.tex
    includes:
      in_header: header.tex

csl: /Users/juanrocha/Documents/styles/annual-reviews.csl
bibliography: refs.bib
documentclass: ar-1col-S2O
header-includes:
- \markboth{Rocha et al.}{Regime shifts and transformations}

# abstract: | 
#    | Current research problems in sustainability science require thinking on nonlinear changes with central questions such as how to avoid them, promote them, or predict them. Regime shifts and transformations literature addresss some of these qeustions, yet there is little cross-fertilization between these two streams of research. Here we review them, propose avenues from future reseach and how together they can be a powerful tool to achieve safe and just futures

# keywords: "resilience, sustainability, transformations, adaptative capacity, regime shifts, adaptive management"
---

```{r setup, include=FALSE}
library(tidyverse)
library(here)

## word counts
## Abstract 150
## Intro 697
## Method 957
## Results 621
## Discussion 1434
## Conclusion 267
## Total main 3976 + 150 abstract
## 
## Targets: Annual Reviews or Environment and Resources
## 9300w, 2 figs, 150 refs (27 pages in published volume)
## required sections:
## - summary points (max 8)
## - Future issues (max 8)
## - Terms and definitions: max 20, definitions of 20 words each

# Outline
# 
# Problem setting [1000w]
# 
# Definitions [1000w, 500w each]
# 
# History of the concepts [1000w]
# 
# Commonalities and differences [500w]
# 
# Current research challenges [1500w]
# 
# Multidimensionality
# 
# Observability and detection
# 
# Management: ethical stakes
# 
# Navigating the space of positive and negative tipping points [1000w]
# 
# Conclusions [500w]
# 
# Key insights (to be updated):
# 
# Clarify the difference and overlaps between transformations and regime shifts theories and the use of concepts such as tipping points
# 
# Highlighting frontiers of research that need both approaches
# 
# Navigating the space of positive and negative tipping points
# 
# \newpage
#

```

\begin{abstract}
Current research problems in sustainability science require thinking on
nonlinear changes with central questions such as how to avoid them,
promote them, or predict them. Regime shifts and transformations
literature addresss some of these qeustions, yet there is little
cross-fertilization between these two streams of research. Here we
review them, propose avenues from future reseach and how together they
can be a powerful tool to achieve safe and just futures
\end{abstract}
\begin{keywords}
resilience, sustainability, transformations, adaptative capacity, regime shifts, adaptive management
\end{keywords}

\maketitle

\tableofcontents

\section{INTRODUCTION}

Sustainability scientists are concerned with nonlinear changes.
Depending on the research problem, the focus could be on avoiding undesirable changes, or promoting desirable ones.
But regardless of direction or desirability, notions of stability and resilience are common ground.
Examples of nonlinear changes include the adoption of new technologies or regulatory frameworks for e.g. the energy transition, the banning of damaging substances for human health or the environment, such as chlorofluorocarbons (CFC) and their impacts on the ozone layer, or the collapse of fish stocks.

There is two strands of sustainability science literature geared to deal with nonlinear changes that share a common origin [@holling1973resilience] but have been diverging in the last couple of decades: regime shifts [@folke2004] and transformations [@olsson2006shooting].
Regime shifts are large, abrupt and persistent changes in the function and structure of social-ecological systems [@biggs2018].
Transformations are about shifting development into new pathways, emergent and often unknown development trajectories [@folke2016].
We expand on definitions, similarities and differences of these two streams of work on the next section.
For now, we want to focus on why this review is timely and what we aim to bring to the larger scientific discussion.

Our review is motivated by the apparent separation of the two streams of work given that both come from the same conceptual framework [@holling1973resilience] but are applied in different ways by different scholarly audiences.
This could result in dimensions that are central on one framework becoming oversimplified in the other.
For example, agency —the ability of actors to influence each other and the environment— is central within transformations framing, but less prominent within regime shifts studies.
Conversely, the role of feedbacks is central in explaining regime shifts, but not always assessed in transformations research.

This lack of dialogue can lead to misuse of terminology, oversimplifciation, jargonification, or simply re-inventing the wheel.
For example, the concept of positive tipping points emphasize desirable transitions such as the adoption of electric vehicles [@lenton2025; @lenton2023global], but omit to analyze which societal groups will benefit or suffer from this transition and what it does imply for social dimensions of sustainability.
Such issues related to justice and power have been studied for three decades in transformations research [@pereira2024equity] and could have highlighted the societal risks entailed in energy transitions.

Sustainability science problems often bring questions of desired for whom, whose voices are heard in decision making, or dimensions of justice with e.g. future generations.
Research in both the transformations and regime shifts strains have already contributed important milestones and tools to deal with such questions.
However, these communities ignoring each other risks missing opportunities or the lack of cross-fertilization between the fields.
In other words, it reduces the probability that we can successfully solve some of the core problems of sustainability science, which by definition, require understanding social and biophysical dynamics simultaneously.

Sustainability science suffers from a proliferation of conceptual frameworks that grows faster than the capacity of scientist to test their validity or adequacy to explain real world problems.
It is not alone, similar scholarly dynamics have been reported in sociology, ecology or physics [@clark2020; @watts2014; @marquet2014].
A diversity of frameworks is an asset when a sustainability scientist treat them as tools to solve research problems.
But it becomes problematic when similar ideas are re-framed over and over under different semantics, yet none of them advance from framework to scientific theory, that is a set of causal statements that produce refutable expectations.

For example, the idea that social and ecological dynamics are intertwined, inseparable or entangled is at least three decades old [@berkes2000linking; @berkes2008navigating; @folke1996; @folke1991; @günther1993; @levin1999fragile], if not older [@renn2020evolution].
In fact, the distinction between the natural and social does not exist in many indigenous cosmologies.
Multiple frameworks have attempted to reshape the idea, from the social-ecological systems framework (SES) [@ostrom2009], socio-ecological technological systems (SETS) [@mcphearson2022; @ramaswami2012], coupled human environmental systems (CHENS) [@liu2007]] and more recently relational approaches [@mancillagarcia2024; @walsh2020], to name a few.
More matured frameworks such as the SES have hundreds of case studies coded supporting the internal and external validity of the framework, and suggesting under which conditions an SES is likely to succeed in sustaining e.g. managing common pool resources [@ostrom2007, e.g. [Library on Governance of Social-Ecological Systems](https://seslibrary.asu.edu)].
Other frameworks lack such rich empirical evidence, although some efforts are on their way in e.g. modeling transformations combining case studies and relational approaches [@schlüter2025].

As we will review, some of the criticisms of regime shifts and transformations are lack of comparative case studies that shed light under which conditions these frameworks have explanatory power, or under which conditions are they likely to fail.
An additional risk of multiple untested frameworks is the use of slightly different wording to mean the same thing (jargonification), for example positive tipping points and transformations.
Here we will review and clarify the nuance on meanings to avoid falling into the jargon trap, or at least make the content more accessible from a pedagogical point of view.

This review aims to outline research frontiers within sustainability science by looking closely to the transformations and regime shifts literature.
We first dive on the definitions and history of these concepts, trying to disentangle what are their common grounds and differences.
We then outline current research challenges and areas where we believe the two strands of work could benefit from working together.
Last, we conclude with a short agenda for future research.

# HISTORY AND DEFINITIONS

While we focus on the phenomena of regime shifts and transformations, the central character of the story is actually the object of study to which these phenomena can happen: social-ecological systems.
Social-ecological systems (SES) are human and natural coupled systems where people interact with biophysical components; they often exhibit nonlinear dynamics, feedbacks, time lags, heterogeneity and resilience [@liu2007a].
Resilience is the ability of a system to withstand disturbances without losing its function, structure and identity [@folke2016; @holling1973resilience].
The resilience scholars have adopted two important evolutionary features: resilience is also the capacity to adapt and transform.
Adaptability is the capacity to learn, combine experience and knowledge to adjust responses, while transformability is the ability to cross thresholds and move the SES into new basins of attraction, into emergent and often unknown development trajectories [@folke2016].

Historically, the conceptualization of resilience has its roots in mathematics.
Holling in his seminal work introducing the concept of resilience [@holling1973resilience] uses bifurcation theory (or catastrophe theory) to suggest minimal cartoons on how ecosystems can change, how they can lose stability and shifts from one configuration to another.
His work would have not been possible without previous developments on bifurcations (e.g. [@Poincare:1885ji; @thom1972stabilité] ).
Several concepts such as the basin of attraction (or attractor), tipping points, thresholds, or stability landscapes have precise definitions within these theories; but sometimes used as metaphors in later conceptualizations.
A basin of attraction, for example, is the region of parameter space that renders the same macroscopic qualitative behavior, it can be an equilibrium point, but also some more complex behaviour such as (a)periodic cycles, long transients, or chaos [@strogatz1994nonlinear].
We will not present formal definitions of these concepts, but instead direct the interested reader to introductory books on the matter [@strogatz1994nonlinear; @sole2011phase; @kuznetsov2023].

In the mathematical context, resilience is the size of the basin of attraction [@holling1973resilience; @ives1995; @Clark1975; @krakovská2023].
Features such as the distance to the boundary, the deepness or slope are often proposed as proxies of resilience [@Clark1975; @Mitra:2015jb; @Menck:2013jq] and enable the development of early warning systems for the proximity to tipping points [@scheffer2009].
Bifurcation theory, and thus notions of resilience, have been successfully applied to fields other than sustainability science and ecology, including but not limited to language, climate systems, the states of matter, cancer dynamics, finance, social dynamics such as poverty traps, believe spreading, voting behavior, segregation, the states of mind, or music, no name a few [@sole2011phase; @scheffer2009critical].
Some of these applications preceded the work of Holling in ecology [@holling1973resilience], and outlines the success of bifurcation theory across the sciences.
However, evolutionary dynamics of adaptations and transformations are not well defined within the seminal work on bifurcations.
While some formalizations have been used to explore for example the evolution of innovations [@kauffman1987; @kauffman2000; @sole2011phase], it remains an open problem to theorize and true generative models that produce agents with behaviors that were not previously coded by the programmer [@kauffman2023].
This problem is common ground to both regime shifts and transformations research.

## Regime shifts

Regime shifts are large, abrupt and persistent changes in the structure and function of social-ecological systems [@folke2004; @biggs2018].
They occur when resilience is eroded and the SES can easily depart from its original configuration and explore different regimes or modes of behavior.
Examples of regime shifts includes the transition from corals to algae dominated reefs [@hughes2017], eutrophication of lakes [@carpenter2020], the shift from forest to savannahs [@hirota2011; @staver2011], the dissapearance of Arctic Sea Ice [@livina2013], the collapse of the thermohaline circulation [@stocker1997; @stommel1961], or the collapse of cooperation [@rand2014; @nowak2006].
Regime shifts are important because they often affect the flow of ecosystem services or nature contributions to people [@carpenter2009], and they present a difficult management problem since they can be hard to predict [@biggs2009; @hastings2010] and sometimes impossible to reverse on time scales relevant for management and policy [@biggs2009; @biggs2018].

Regime shifts are typically characterized by their feedbacks, tipping points, thresholds, and hysteresis.
Feedbacks are mutually interacting processes that can act to amplify or damp its response to a force [@roe2009].
For example, the moisture recycling feedback is responsible for the maintenance of tropical rain forest like the Amazon, the more forest area there are, the more humidity is transpired, which in turn feeds the clouds in the atmosphere that rains back to support the water demands of the forest [@zemp2017].
In drylands, facilitation feedbacks through shading, soil moisture and nutrient retention are responsible for shifts between grasslands, shrublands and deserts [@maestre2016; @berdugo2022].
In coral reefs, herbivory plays a fundamental role in mediating the shift between corals and macro-algae [@hughes2007].
Or in social systems, poverty traps are feedbacks mediated by income, education, or disease that keeps people in poverty despite development efforts [@bowles2016poverty; @banerjee2012poor].
Balancing feedbacks are responsible for the stabilization of a particular regime, while reinforcing feedbacks are responsible for out of equilibrium dynamics [@biggs2018].
Thus weakening or strenghtening of feedbacks can trigger nonlinear transitions.

Tipping points are critical parameter values at which a system change its long term qualitative behavior.
They are bifurcation points or sigularities [@strogatz1994nonlinear; @thom1972stabilité].
Speaking of a point implies the assumption of a one-dimensional system with one co-dimension, that is a very simplified version of reality where there is only one state variable that we care about (e.g. one species) and one driver affecting it (e.g. temperature).
In reality, social-ecological system are complex adaptive systems composed by multiple interacting entities (e.g. species, businesses, cities) that are driven by multiple interacting forces that are not necessarily reducible to one (co)dimension [@levin1999fragile; @rocha2015].
When thinking of tipping phenomena in real life contexts, one should think on tipping lines or an hypersurface, that is all the combination of critical parameters of all potential drivers and state variables, rather than points.
Tipping points and thresholds are often used interchangeably.
However, in mathematics they have a slightly different meaning, the threshold is the separatrix between two basins of attraction [@strogatz1994nonlinear].

Observing and measuring tipping points or thresholds in real life context is very difficult, and for that reason sometimes deemed not useful beyond theoretical work [@hillebrand2020; @kopp2024].
However, scientists have been able to identify some tipping points with different degrees of certainty.
For example, hypoxia and anoxia are regime shifts common in coastal areas where dissolved oxygen fall under levels that are not tolerable for different groups of organisms.
Levels below $2mlO_{2}/L$ gives rise to hypoxia a condition where many fish die-off, and below $0.5mlO_{2}/L$ to annoxia a condition that is toxic for humans [@diaz2008; @breitburg2018].
Tipping points in precipitation have been proposed for the transitions from forest to savanna ($1500mm$), savanna to forest ($2000mm$), and savanna to deserts ($<500mm$) [@hirota2011], which have been confirmed in observational studies [@staver2011; @aleman2020; @verbesselt2016].
Similarly, global thresholds for aridity has been reported for dryland ecosystems (0.54, 0.7 and 0.8) [@berdugo2020], while thresholds for deforestation in rain forest have been proposed (40% for the Amazon, 25% under precautionary principle, currently at 20%) [@nobre2016; @lovejoy2018].
Several Earth subsystems can have tipping points driven by climate change.
A recent expert elicitation speculates that six of these subsystems are highly likely to experience tipping dynamics under $2^\circ C$ of warming, with additional four tipping under $4^\circ C$ of warming [@armstrongmckay2022].

Hysteresis is the reason why some regime shifts are really hard, sometimes impossible to reverse.
Hysteresis means that tipping point in one direction of the shift is different from the tipping point in the opposite direction.

\begin{marginnote}[]
\entry{Feedback}{Mutually interacting processes that can amplify or damp its response}
\entry{Tipping point}{Critical parameter values at which a system change its long term qualitative behavior}
\entry{Hysteresis}{The critical point in one direction of the shift is different from the critical point in the opposite direction}
\end{marginnote}

-   features: feedbacks, tipping points, hysteresis,

-   jargon: phase transitions, critical transitions, phase shifts, bifurcations.

-   adaptive management, co-management

-   detection (frontiers of research?)

-   limitations: evolutionary dynamics, empirical evidence, not fully social-ecological.

-   misconceptions: stable states, necessity of hysteresis.

## Transformations

## Similarities

## Differences

# RESEARCH CHALLENGES

# CONCLUSIONS

\begin{summary}[SUMMARY POINTS]
\begin{enumerate}
\item Summary point 1. These should be full sentences.
\item Summary point 2. These should be full sentences.
\item Summary point 3. These should be full sentences.
\item Summary point 4. These should be full sentences.
\end{enumerate}
\end{summary}

\begin{issues}[FUTURE ISSUES]
\begin{enumerate}
\item Future issue 1. These should be full sentences.
\item Future issue 2. These should be full sentences.
\item Future issue 3. These should be full sentences.
\item Future issue 4. These should be full sentences.
\end{enumerate}
\end{issues}

\section*{DISCLOSURE STATEMENT}

If the authors have noting to disclose, the following statement will be used: The authors are not aware of any affiliations, memberships, funding, or financial holdings that might be perceived as affecting the objectivity of this review.

\section*{ACKNOWLEDGMENTS}

Acknowledgements, general annotations, funding.

\section*{LITERATURE CITED}